Population structure and phenotypic variation of *Sclerotinia sclerotiorum* from dry bean in the United States

### 2017-06-30 doc/manuscript/manuscript.Rmd
update manuscript after discussion with Syd
Still needs tweaking

---
title: Population structure and phenotypic variation of *Sclerotinia sclerotiorum* from dry bean throughout the United States, and from Mexico, France, and Australia
preprint: false
author:
  - name: Zhian N. Kamvar
    email: zkamvar@unl.edu
    affiliation: 1
  - name: Bimal Sajeewa Amadarasa
    email: bamadarasa2@unl.edu
    affiliation: 1
  - name: R. Jhala
    email: rjhala2@unl.edu
    affiliation: 1
  - name: James Steadman
    email: jsteadman1@unl.edu
    affiliation: 1
  - name: Sydney E. Everhart
    email: everhart@unl.edu
    affiliation: 1
    corresponding: true
affiliation:
  - code: 1
    address: Department of Plant Pathology, University of Nebraska-Lincoln, Lincoln 68583
abstract: >
  `r paste(readLines("abstract.md"), collapse = "\n  ")`
header-includes: >
  \usepackage{indentfirst}
bibliography: ssc_bibliography.bib
output: 
  rticles::peerj_article:
    citation_package: natbib
    base_format: bookdown::pdf_document2 # for using \@ref()
---

@@

<!-- Most of the previous studies on *S. sclerotiorum* had been conducted in the USA, Canada and Australia but not in Europe. 
In those studies, different dry bean cultivars had been rarely used as  hosts to investigate any host-pathogen relationships or specificity.
No white mold resistant dry bean cultivars exist.
Canopy architecture and growth pattern of some cultivars enable improved disease avoidance [@kolkman2002agronomic]. -->
To date, no commercial white mold-resistant dry bean cultivars exist. 
Field trials in some regions of the US have shown the G122 cultivar is potentially resistant while the Beryl cultivar is susceptible (pers. comm. with J. Steadman).
Understanding genetic relationships and reproduction behavior of *S. sclerotiorum* populations from locations worldwide is beneficial for plant breeders seeking to develop white mold resistance cultivars for deployment worldwide. 
This information is also helpful in making disease management decisions and developing disease resistance screening methods.

@@

We analyzed *S. sclerotiorum* isolates collected from dry bean cultivars from 2003 to 2012 from different geographic locations in America, Europe and Australia using microsatellite markers to delineate genetic differences and reproduction behavior.
The specific aims of our study were:  a) To assess association of phenotypic data (virulence) and environmental data (bean cultivar, region, field, year) on genotypic variation of *S. sclerotiorum* isolates, b) To test the hypothesis that *S. sclerotiorum* populations of dry beans originated from different continents were genetically different, and c) although previous studies show that *S. sclerotiorum* populations are largely clonal, few compare populations from different continents [@lehner2017sclerotinia], c) To identify localized genotypes representative of each location and also investigate if there are widespread genotypes common among locations. d) To test potential for sexual reproduction in sub populations of the pathogen isolates under different conditions such as year or field in each region, year/region, field/region/year, and e) Test if pathogen within-field populations differentiate over time.

@@

Mycelial growth of each isolate was generated by activating sclerotia according to the method described previously [@matheron1989in-vitro].
These isolates (N = 366) were phenotypically characterized by determining mycelial compatibility group MCG and virulence.
MCG was determined through co-culturing isolates in pairs on PDA plates, where incompatibility of different MCG resulted in formation of a mycelial barrage [@kohn1990mycelial] and isolates belonging to the same MCG did not form a mycelial barrage.
Virulence of each isolate was assessed using a straw test [@petzoldt1996straw; @teran2006modified] that rates necrotic lesion on a scale of 1-9, with ascending numbers assigned for increasing aggressiveness.
DNA was purified from mycelium of each isolate using a Qiagen DNeasy Plant Mini Kit (Qiagen Inc., Valencia, CA) according to the manufacturer's instructions.

@@

### Virulence Assessment {-}

@@

Genotypic evenness [@grunwald2003analysis], which is a measure of distribution of genotype abundances and unbiased gene diversity ($H_{exp}$) [@nei1978estimation] were calculated using *poppr* package in R [@kamvar2014poppr].
Microsatellite data were also tested for evidence of linkage disequilibrium, where significant disequilibrium between pairs of loci in a population is attributed to linkage between loci and may be an indication of clonal reproduction if loci are physically unlinked, whereas linkage equilibrium can be an indication of outcrossing populations.
The linkage disequilibrium of clone-corrected data was determined by calculating the index of association ($I_{A}$) and its standardized version ($\bar{r}_d$) and a permutation approach with 999 permutations was used to test whether observed values were significantly different from data randomized at each locus [@agapow2001indices; @brown1980multilocus].
The value of $I_A$ tends to increase with the number of loci and therefore we made our inferences based on $\bar{r}_d$, which standardizes the number of loci tested and considered a more robust method [@agapow2001indices].
Higher observed $\bar{r}_d$ or $I_A$ values would indicate clonally reproducing populations.

@@

@@

## Virulence {-}

The MCGs with the highest frequency were MCG 5 and MCG 44 representing 73 and 44 isolates, respectively **(Fig. 4)**. <!-- Maybe add aggressiveness information here. -->
MCGs that were unique to a region are shown in Fig. 5 while MCGs that represented at least 2 or more regions are in Fig 6. 20 MCGs were common in two or more regions and the rest were distributed uniquely in different regions.
ID, MN, NE, NY, and WI did not have any unique/isolated MCGs.
The virulence of isolates ranged from 1.4 to 7.9 with a mean of 5.01.
There were six isolates that fell into upper outlier category (virulence > 7.6) and one isolate in the lower outlier range (virulence < 2.3) (box plot results not shown).
The mean virulence of the 10 most abundant MCGs ranged from 4.6 (MCG 49) to 6.02 (MCG 44) (see Fig 4 for 10 most abundant MCGs).
ANOVA of these 10 MCGs indicated significant difference for virulence levels at P <0.0001.
Mean separation of 10 most common MCGs showed MCG 44 is different from MCGs 9, 1, 45, 4, 53, and 49 while the virulence levels of MCGs 2, 5, 3 failed to show any difference to MCG 44 or the other MCGs.

The number of observed alleles per locus ranged anywhere from 2 to 10 (Table \@ref(tab:locus-stats)).
All loci were informative at cut off level of 2/N where N was the total number of individuals.
The genetic diversity indices of *S. sclerotiorum* populations from different states of the USA, France, Mexico and Australia based on SSR analysis are given in Table 5.
The 165 unique MLGs identified among all 366 isolates are given in Supplementary Table 1.
Gene diversity ($H_{exp}$) of populations from different regions ranged from 0.259 to 0.596 **(Table 5)**.
Evenness of genotypes of all regions was high, with the exception of ND, where a single genotype was represented by 20 isolates.
There was one private allele in the WA population (204 bp allele of 55-4 locus) and two private alleles in the ND population (two 366 bp alleles in 9-2 locus and two 220 bp alleles in 12-2 locus).

----------------------------------------------
Locus  Range    No. alleles $H_{exp}$ Evenness 
------ ------- ------------ --------- --------
5-2    318-324            4      0.45     0.62

@@

*mean* 153-495         6.27      0.58     0.69

----------------------------------------------

Table: (\#tab:locus-stats) Allelic diversity on full data set at loci used in this study. $H_{exp}$ = Nei's Gene Diversity [@nei1978estimation].

@@

The dbRDA was performed on clone-corrected isolates (N = 318) after grouping them by location, host, region, and year.
The forward-backward selection process revealed severity, year, region, and host to be the optimal variables for the reduced model, accounting for 17.6% of the variation.
ANOVA revealed that the reduced model was significant with an adjusted R^2^ of 0.032 (P = 0.001).
The marginal effects of state, host, and year were also significant (P = 0.001), but virulence was not significant (P = 0.045).
Much of the variation appears to be driven by isolates from Mexico (Fig. \@ref(rda-biplot)) and 2005, but it's worth noting that all Mexican isolates were collected in 2005.
Variance partitioning of the independent variables indicated virulence to be the least influential factor with 0.1% contributing to explaining the variation of molecular data whereas the combination of variables accounted for 3.3%.

@@

Our study was conducted to characterize *S. sclerotiorum* isolates from different geographic areas to determine important phenotypic variables influencing genotypes; to ascertain reproductive behavior of the pathogen; to test whether different sampling years or locations have any effect on genotype/sexuality; and to identify abundant and unique genotypes across isolates and geographies.
The dbRDA revealed region, year, and host as the most influential phenotypic variables on genotypes.
However, further molecular analysis of pathogen isolates from selected hosts did not show any significant genetic difference among populations.
According to dbRDA, virulence and location of isolates had little or no effect.
Previous studies on *S. sclerotiorum* from alfalfa soil [@attanayake2012genetic] and potato [@atallah2004high], and canola [@attanayake2013sclerotinia] plants reported little or no correlation of genotypes on pa25thogen virulence.
This is important information for resistant breeding since genotypes of *S. sclerotiorum* isolates seem to depend to a certain degree on different levels of region and also on environmental conditions of the sampling year.
For instance our analysis indicated that MX, 2009, and 2005 were highly influential predictors of *S. sclerotiorum* genotypes (Fig. 7).
Conversely, France and Australia, which are geographically distant to North America, were not within the top 10 predictors of genetic composition of isolates (Fig. 7).
According to dbRDA, Mexican region imposed the highest impact on microsatellite loci of *S. sclerotiorum* isolates.
We used Bruvo's distance for calculating genetic distances of different isolates since it incorporates stepwise mutation model and applicable to microsatellite analysis.

@@

Interestingly, all Mexican genotypes were unique to that region.
On the other hand CA genotypes were present in the West Coast states of OR and WA which may be due to movement of sclerotia with planting material. <!-- ZNK: Do we have some sort of evidence for this, or is this pure speculation? -->
<!-- 
I'm not sure what the point of this paragraph is.
I believe I can put the discussion of the graph results here. 
-->
Surprisingly, France and Australian genotypes were also isolated in the USA.
MLG 4 (see supplemental Table 1) was recorded from France, Australia and MN while France and USA isolates had four common MLGs.
Similarly, five MLGs were recorded common to both Australia and USA isolates.
However, there were six unique genotypes in clone-corrected France population (N= 6/14) and three genotypes in CA population (N = 3/15).
MLGs from Australian region, our samples were limited to 6, which is not a sufficient number to make inferences.
Our results indicated that different continents were not a defining factor for microsatellite genotypes of *S. sclerotiorum* isolated from dry bean.
This may be due to involuntary movement of sclerotia by human activities and probably because of recent speciation of *S. sclerotiorum*. 
A phylogenetic study on Sclerotinaceae showed that most of the sclerotial lineages including different species of Sclerotinia had evolved relatively recently from a common ancestor [@carbone1993ribosomal].
Few multilocus genotypes were unique to each sampling state of the USA but most of them were common among several states (Supplemental Table 1).
For instance, Michigan state from which we had the largest collection of isolates (N = 62) consisted of 43 MLGs.
Of these, 32 MLGs were present in at least one of the 5 states of NE, ND, WA, CO, MN; 4 MLGs were present in France isolates; and 11 were unique to MI sites (see Supplemental Table 1 for unique MLGs in different regions).
When selecting *S. sclerotiorum* isolates for screening dry bean germplasm for disease resistance, plant breeders may find above information useful.
A study on genetic diversity of *S. sclerotiorum* from multiple host crops in the North central region of the USA revealed there was little evidence for genetic isolation of this pathogen by host crop or by distance [@aldrich-wolfe2015genetic].
Though our isolates didn't come from different crops, we had isolates from many resistant and susceptible dry bean cultivars, which also didn't show genetic differentiation according to AMOVA.
It seems Mexican pathogen isolates are different from the rest and warrants further investigation using a larger population.
Some France genotypes were common in USA isolates but at the same time 43% (6/14) of the France genotypes were unique to that region indicating white mold resistant cultivars produced in the USA might not work in that region. <!-- ZNK: I'm not entirely sold on this conclusion. -->
Geographic distance between different states of the USA seem not a barrier for gene transfer of *S. sclerotiorum* isolates though unique genotypes did prevail at a low frequency in each state. <!-- ZNK: I don't know that this adds anything. -->
The genetic diversity of regions with >10 isolates had similar values (0.43 - 0.59) except for Mexican isolates.
$H_{exp}$ of MX was 0.26 and indicated less exposure of sampling sites to outside inocula. <!-- ZNK: Does this mean that there's heavy inbreeding here? -->
Test for linkage disequilibrium showed evidence of clonal reproduction for MX isolates (Ia = 2.61, P.Ia = 0.001, $\bar{r}_d$ = 0.37, P.$\bar{r}_d$  = 0.001).
This may have been another reason for low gene diversity.

@@

>>>>

---
title: Population structure and phenotypic variation of *Sclerotinia sclerotiorum* from dry bean in the United States
preprint: false
author:
  - name: Zhian N. Kamvar
    email: zkamvar@unl.edu
    affiliation: 1
  - name: Bimal Sajeewa Amadarasa
    email: bamadarasa2@unl.edu
    affiliation: 1
  - name: Rachana Jhala
    email: rjhala2@unl.edu
    affiliation: 1
  - name: James Steadman
    email: jsteadman1@unl.edu
    affiliation: 1
  - name: Sydney E. Everhart
    email: everhart@unl.edu
    affiliation: 1
    corresponding: true
affiliation:
  - code: 1
    address: Department of Plant Pathology, University of Nebraska-Lincoln, Lincoln 68583
abstract: >
  `r paste(readLines("abstract.md"), collapse = "\n  ")`
header-includes: >
  \usepackage{indentfirst}
bibliography: ssc_bibliography.bib
output: 
  rticles::peerj_article:
    citation_package: natbib
    base_format: bookdown::pdf_document2 # for using \@ref()
---

@@

<!-- Most of the previous studies on *S. sclerotiorum* had been conducted in the USA, Canada and Australia but not in Europe. 
In those studies, different dry bean cultivars had been rarely used as  hosts to investigate any host-pathogen relationships or specificity.
No white mold resistant dry bean cultivars exist.
Canopy architecture and growth pattern of some cultivars enable improved disease avoidance [@kolkman2002agronomic]. -->
To date, no commercial white mold-resistant dry bean cultivars exist. 
Field trials in some regions of the US have shown the G122 cultivar is potentially resistant while the Beryl cultivar is susceptible (pers. comm. with J. Steadman).
<!-- SEE: Evaluating cultivars, there should be some BIC reports that we can cite; personal communication is kind of eh. -->
Understanding genetic relationships and reproduction behavior of *S. sclerotiorum* populations from locations worldwide is beneficial for plant breeders seeking to develop white mold resistance cultivars for deployment worldwide. 
This information is also helpful in making disease management decisions and developing disease resistance screening methods.

@@

We analyzed *S. sclerotiorum* isolates collected from dry bean cultivars from 2003 to 2012 from different geographic locations in America, Europe and Australia using microsatellite markers to delineate genetic differences and reproduction behavior.
The specific aims of our study were:  a) To assess association of phenotypic data (aggressiveness) and environmental data (bean cultivar, region, field, year) on genotypic variation of *S. sclerotiorum* isolates, b) To test the hypothesis that *S. sclerotiorum* populations of dry beans originated from different continents were genetically different, and c) although previous studies show that *S. sclerotiorum* populations are largely clonal, few compare populations from different continents [@lehner2017sclerotinia], c) To identify localized genotypes representative of each location and also investigate if there are widespread genotypes common among locations. d) To test potential for sexual reproduction in sub populations of the pathogen isolates under different conditions such as year or field in each region, year/region, field/region/year, and e) Test if pathogen within-field populations differentiate over time.

@@

Mycelial growth of each isolate was generated by activating sclerotia according to the method described previously [@matheron1989in-vitro].
These isolates (N = 366) were phenotypically characterized by determining mycelial compatibility group MCG and aggressiveness.
MCG was determined through co-culturing isolates in pairs on PDA plates, where incompatibility of different MCG resulted in formation of a mycelial barrage [@kohn1990mycelial] and isolates belonging to the same MCG did not form a mycelial barrage.
Aggressiveness of each isolate was assessed using a straw test [@petzoldt1996straw; @teran2006modified] that rates necrotic lesion on a scale of 1-9, with ascending numbers assigned for increasing aggressiveness.
DNA was purified from mycelium of each isolate using a Qiagen DNeasy Plant Mini Kit (Qiagen Inc., Valencia, CA) according to the manufacturer's instructions.

@@

### Aggressiveness Assessment {-}

@@

Genotypic evenness [@grunwald2003analysis], which is a measure of distribution of genotype abundances and unbiased gene diversity (*h*) [@nei1978estimation] were calculated using *poppr* package in R [@kamvar2014poppr].
Microsatellite data were also tested for evidence of linkage disequilibrium, where significant disequilibrium between pairs of loci in a population is attributed to linkage between loci and may be an indication of clonal reproduction if loci are physically unlinked, whereas linkage equilibrium can be an indication of outcrossing populations.
The linkage disequilibrium of clone-corrected data was determined by calculating the index of association ($I_{A}$) and its standardized version ($\bar{r}_d$) and a permutation approach with 999 permutations was used to test whether observed values were significantly different from data randomized at each locus [@agapow2001indices; @brown1980multilocus].
The value of $I_A$ tends to increase with the number of loci and therefore we made our inferences based on $\bar{r}_d$, which standardizes the number of loci tested and considered a more robust method [@agapow2001indices].
Higher observed $\bar{r}_d$ or $I_A$ values would indicate clonally reproducing populations.

@@

The associations between MCGs and MLGs are complicated \@ref(biggraph)

@@

## Aggressiveness {-}

The MCGs with the highest frequency were MCG 5 and MCG 44 representing 73 and 44 isolates, respectively **(Fig. 4)**. <!-- Maybe add aggressiveness information here. -->
MCGs that were unique to a region are shown in Fig. 5 while MCGs that represented at least 2 or more regions are in Fig 6. 20 MCGs were common in two or more regions and the rest were distributed uniquely in different regions.
ID, MN, NE, NY, and WI did not have any unique/isolated MCGs.
The aggressiveness of isolates ranged from 1.4 to 7.9 with a mean of 5.01.
There were six isolates that fell into upper outlier category (aggressiveness > 7.6) and one isolate in the lower outlier range (aggressiveness < 2.3) (box plot results not shown).
The mean aggressiveness of the 10 most abundant MCGs ranged from 4.6 (MCG 49) to 6.02 (MCG 44) (see Fig 4 for 10 most abundant MCGs).
ANOVA of these 10 MCGs indicated significant difference for aggressiveness levels at P <0.0001.
Mean separation of 10 most common MCGs showed MCG 44 is different from MCGs 9, 1, 45, 4, 53, and 49 while the aggressiveness levels of MCGs 2, 5, 3 failed to show any difference to MCG 44 or the other MCGs.

The number of observed alleles per locus ranged anywhere from 2 to 10 (Table \@ref(tab:locus-stats)).
All loci were informative at cut off level of 2/N where N was the total number of individuals.
The genetic diversity indices of *S. sclerotiorum* populations from different states of the USA, France, Mexico and Australia based on SSR analysis are given in Table 5.
The 165 unique MLGs identified among all 366 isolates are given in Supplementary Table 1.
Gene diversity (*h*) of populations from different regions ranged from 0.259 to 0.596 **(Table 5)**.
Evenness of genotypes of all regions was high, with the exception of ND, where a single genotype was represented by 20 isolates.
There was one private allele in the WA population (204 bp allele of 55-4 locus) and two private alleles in the ND population (two 366 bp alleles in 9-2 locus and two 220 bp alleles in 12-2 locus).

--------------------------------------------
Locus  Range    No. alleles    *h*  Evenness 
------ ------- ------------ ------ ---------
5-2    318-324            4   0.45      0.62

@@

mean   153-495         6.27   0.58      0.69

--------------------------------------------

Table: (\#tab:locus-stats) Allelic diversity on full data set at loci used in this study. *h* = Nei's Gene Diversity [@nei1978estimation].

@@

The dbRDA was performed on clone-corrected isolates (N = 318) after grouping them by location, host, region, and year.
The forward-backward selection process revealed severity, year, region, and host to be the optimal variables for the reduced model, accounting for 17.6% of the variation.
ANOVA revealed that the reduced model was significant with an adjusted R^2^ of 0.032 (P = 0.001).
The marginal effects of state, host, and year were also significant (P = 0.001), but aggressiveness was not significant (P = 0.045).
Much of the variation appears to be driven by isolates from Mexico (Fig. \@ref(rda-biplot)) and 2005, but it's worth noting that all Mexican isolates were collected in 2005.
Variance partitioning of the independent variables indicated aggressiveness to be the least influential factor with 0.1% contributing to explaining the variation of molecular data whereas the combination of variables accounted for 3.3%.

@@

Our study was conducted to characterize *S. sclerotiorum* isolates from different geographic areas to determine important phenotypic variables influencing genotypes; to ascertain reproductive behavior of the pathogen; to test whether different sampling years or locations have any effect on genotype/sexuality; and to identify abundant and unique genotypes across isolates and geographies.
The dbRDA revealed region, year, and host as the most influential phenotypic variables on genotypes.
However, further molecular analysis of pathogen isolates from selected hosts did not show any significant genetic difference among populations.
According to dbRDA, aggressiveness and location of isolates had little or no effect.
Previous studies on *S. sclerotiorum* from alfalfa soil [@attanayake2012genetic] and potato [@atallah2004high], and canola [@attanayake2013sclerotinia] plants reported little or no correlation of genotypes on pa25thogen aggressiveness.
This is important information for resistant breeding since genotypes of *S. sclerotiorum* isolates seem to depend to a certain degree on different levels of region and also on environmental conditions of the sampling year.
For instance our analysis indicated that MX, 2009, and 2005 were highly influential predictors of *S. sclerotiorum* genotypes (Fig. 7).
Conversely, France and Australia, which are geographically distant to North America, were not within the top 10 predictors of genetic composition of isolates (Fig. 7).
According to dbRDA, Mexican region imposed the highest impact on microsatellite loci of *S. sclerotiorum* isolates.
We used Bruvo's distance for calculating genetic distances of different isolates since it incorporates stepwise mutation model and applicable to microsatellite analysis.

@@

Interestingly, all Mexican genotypes were unique to that region.
On the other hand CA genotypes were present in the West Coast states of OR and WA which may be due to movement of sclerotia with planting material. <!-- ZNK: Do we have some sort of evidence for this, or is this pure speculation? -->
<!-- 
I'm not sure what the point of this paragraph is.
I believe I can put the discussion of the graph results here. 
-->
Surprisingly, France and Australian genotypes were also isolated in the USA.
MLG 4 (see supplemental Table 1) was recorded from France, Australia and MN while France and USA isolates had four common MLGs.
Similarly, five MLGs were recorded common to both Australia and USA isolates.
However, there were six unique genotypes in clone-corrected France population (N= 6/14) and three genotypes in CA population (N = 3/15).
MLGs from Australian region, our samples were limited to 6, which is not a sufficient number to make inferences.
Our results indicated that different continents were not a defining factor for microsatellite genotypes of *S. sclerotiorum* isolated from dry bean.
This may be due to involuntary movement of sclerotia by human activities and probably because of recent speciation of *S. sclerotiorum*. 
A phylogenetic study on Sclerotinaceae showed that most of the sclerotial lineages including different species of Sclerotinia had evolved relatively recently from a common ancestor [@carbone1993ribosomal].
Few multilocus genotypes were unique to each sampling state of the USA but most of them were common among several states (Supplemental Table 1).
For instance, Michigan state from which we had the largest collection of isolates (N = 62) consisted of 43 MLGs.
Of these, 32 MLGs were present in at least one of the 5 states of NE, ND, WA, CO, MN; 4 MLGs were present in France isolates; and 11 were unique to MI sites (see Supplemental Table 1 for unique MLGs in different regions).
When selecting *S. sclerotiorum* isolates for screening dry bean germplasm for disease resistance, plant breeders may find above information useful.
A study on genetic diversity of *S. sclerotiorum* from multiple host crops in the North central region of the USA revealed there was little evidence for genetic isolation of this pathogen by host crop or by distance [@aldrich-wolfe2015genetic].
Though our isolates didn't come from different crops, we had isolates from many resistant and susceptible dry bean cultivars, which also didn't show genetic differentiation according to AMOVA.
It seems Mexican pathogen isolates are different from the rest and warrants further investigation using a larger population.
Some France genotypes were common in USA isolates but at the same time 43% (6/14) of the France genotypes were unique to that region indicating white mold resistant cultivars produced in the USA might not work in that region. <!-- ZNK: I'm not entirely sold on this conclusion. -->
Geographic distance between different states of the USA seem not a barrier for gene transfer of *S. sclerotiorum* isolates though unique genotypes did prevail at a low frequency in each state. <!-- ZNK: I don't know that this adds anything. -->
The genetic diversity of regions with >10 isolates had similar values (0.43 - 0.59) except for Mexican isolates.
*h* of MX was 0.26 and indicated less exposure of sampling sites to outside inocula. <!-- ZNK: Does this mean that there's heavy inbreeding here? -->
Test for linkage disequilibrium showed evidence of clonal reproduction for MX isolates (Ia = 2.61, P.Ia = 0.001, $\bar{r}_d$ = 0.37, P.$\bar{r}_d$  = 0.001).
This may have been another reason for low gene diversity.

@@

# Supplementary Information {-}
\newpage

<!-- This part is necessary for adding the "S" before supplementary figures.
Will not work with HTML outputs, though. -->

\setcounter{table}{0}
\renewcommand{\thetable}{S\arabic{table}}
\setcounter{figure}{0}
\renewcommand{\thefigure}{S\arabic{figure}}

![
Graph showing associations between Mycelial Compatibility Groups dotted outline and Multilocus Genotypes full outline
](../../results/figures/publication/FigureS2.pdf){#biggraph width=100%}

\newpage
